---
title: "us_tilapia_imports.Rmd"
author: "Jared Petry"
date: "2022-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(janitor) 
library(here) 
library(tidyverse) 
```

Task 2 

```{r}
us_tilapia_imports <- read_csv(here::here("data", "us_tilapia_imports.csv")) 

summary(us_tilapia_imports)
names(us_tilapia_imports) #gives the names of all the columns 
dim(us_tilapia_imports) 
```

```{r}
#check classes of columns
class(us_tilapia_imports$country)
class(us_tilapia_imports$`2014`)
#so the country is characters and the years are numeric 
```

Now clean up the data into tidy format 

- Use pivot_longer() to reshape the data into long format (currently, the variable “year” is spread across multiple columns). Remember to store the output so you will be able to use the reshaped data this creates.

```{r}
my_tilapia <- us_tilapia_imports |> 
  pivot_longer(cols = -country, names_to = "year", values_to = "imports")

my_tilapia_numeric <- my_tilapia |> mutate(year = as.numeric(year))
```

Use dplyr::group_by() %>% summarize() to find the total US tilapia imports by year, store as yearly_tilapia_tot

```{r}
yearly_tilapia_tot <- my_tilapia_numeric |> 
  group_by(year) |> 
  summarize(total_imports = sum(imports, na.rm = TRUE))

yearly_tilapia_tot 
```

Make a graph of the above 
```{r}
ggplot(data = yearly_tilapia_tot, aes(x = year, y = total_imports)) +
  geom_line() 

#ggsave(here::here("figs", "yearly_tilapia_tot.png"))
```

```{r}
sub_tilapia <- my_tilapia_numeric |> 
  filter(country %in% c("Ecuador", "Honduras", "Costa Rica", "Mexico")) |> 
  na.omit()
```

graph it 

```{r}
ggplot(data = sub_tilapia, aes(x = year, y = imports, color = country)) + 
  geom_point() + 
  labs(title = "yessir thats tilapia", x = "yeabud", y = "amount tilapia")
```

